---
title: "Equotip hardness measurements - raw material comparison"
author: "Lisa Schunk"
date: "`r Sys.Date()`"
output:
    html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---


```{r Knitr Options, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, indent="", error=TRUE)

```

---


# Goal of the script
This script reads the xlsx file (measurements have been generated with the Equotip Leeb C rebound)  and formats the data for a statistical analysis.  
The script will:

1. Read in the original xlsx file and organise the data 
2. Plot the data 
3. Write an XLSX-file and save an R object ready for further analysis in R 



```{r}
dir_in <- "analysis/raw_data/"
dir_out <- "analysis/plots"

```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.
The knit directory for this script is the project directory.


---

# Load packages
```{r Libraries}
pack_to_load <- c("tidyverse", "readr", "openxlsx", "tools", "ggplot2", "readxl")
sapply(pack_to_load, library, character.only = TRUE, logical.return = TRUE)

```


---


# Get name, path and information of the file
```{r}
data_file <- list.files(dir_in, pattern = "\\.xlsx$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(files = basename(names(md5_in)), checksum = md5_in, 
                  row.names = NULL)

```


# Read in original xlsx-file
```{r}
imp_data <- read.xlsx(xlsxFile = data_file, sheet = 1, startRow = 3, colNames = TRUE, 
                      rowNames = FALSE, skipEmptyRows = FALSE) 

imp_data$hardness.in.HLC <- as.numeric(imp_data$hardness.in.HLC)

# reorder raw.material column 
imp_data[["raw.material"]] <- factor(imp_data[["raw.material"]], levels = c("flint","lydite", "obsidian", "quarzite", "grogged_porcelain", "ungrogged_porcelain", "stone_ware"))
str(imp_data)


flint <- imp_data[1:32,]
lydite <- imp_data[33:62,]
grogged_porcelain <- imp_data[63:87,]
ungrogged_porcelain <- imp_data[88:115,]
stone_ware <- imp_data[116:141,]
quarzite <- imp_data[142:164,]
obsidian <- imp_data[165:194,]


```


# Data analysis - stats  
```{r}
# descriptive statistics 
# flint
length(flint[["ID"]])
summary(flint[["hardness.in.HLC"]])

# schist
length(lydite[["ID"]])
summary(lydite[["hardness.in.HLC"]])

# grogged_porcelain
length(grogged_porcelain[["ID"]])
summary(grogged_porcelain[["hardness.in.HLC"]])

# ungrogged_porcelain
length(ungrogged_porcelain[["ID"]])
summary(ungrogged_porcelain[["hardness.in.HLC"]])

# stone_ware
length(stone_ware[["ID"]])
summary(stone_ware[["hardness.in.HLC"]])

# quarzite
length(quarzite[["ID"]])
summary(quarzite[["hardness.in.HLC"]])

# obsidian
length(obsidian[["ID"]])
summary(obsidian[["hardness.in.HLC"]])

# create new data frame with mean value 
stats <- imp_data %>% group_by(raw.material) %>% summarize(m = mean(hardness.in.HLC))
# change number of decimal digits 
stats$m <- round(stats$m, 1)


```


# Data analsysis - plot 
```{r}
# boxplot
data_plot <- ggplot (imp_data, aes(y = hardness.in.HLC, x = raw.material)) + 
             theme_classic() + 
             theme(legend.title = element_blank()) +
             geom_boxplot(fill = c("#52854c", "#985633", "#134680", "#ffdb6d", "#9a0f0f", "#c3d7a4", "#487bb6")) +
             geom_jitter() + 
             geom_text(data = stats, aes(y = m, label = m),  colour = 'black', size = 4, vjust = 13.1) +
             labs(x = "", y = "Leeb Rebound Hardness in HLC", title = "") +
             scale_x_discrete(labels = c("flint", "lydite", "obsidian", "quarzite", "gr. porcelain", 
                                         "ungro. porcelain", "stone ware")) +
             guides(fill = "none")

print(data_plot)


```


# Save data
## Define output file name
```{r}
file_out <- paste0(basename(file_path_sans_ext(data_file)), ".pdf")
ggsave(filename = file_out, plot = data_plot, path = dir_out, device = "pdf")

```
The files will be saved as "`r paste0("~/", dir_out, ".[ext]")`".


---

# sessionInfo() and RStudio version

```{r}
sessionInfo()

```

RStudio version `r readLines("RStudioVersion.txt", n = 1)`.



# Cite R packages used
```{r Citation}
for (i in pack_to_load) print(citation(i), bibtex = FALSE)

```

--- 

END OF SCRIPT